Palomar 5 Gaps

### 2024-09-23 paper/main.tex
touched up detections

    \subsubsection*{Density Maps and Profiles}
      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \textit{full} simulations using the tail coordinate system. We inspect these differences by generating 2D density maps of the streams. Additionally, we marginalize over the $y'$-coordinate to produce 1D density profiles along the $x'$-axis. These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).

      \begin{figure*}
        \centering
        \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
        \caption{A comparison between the density maps and profiles of the \textit{full} and \textit{vanilla} simulations is presented. The locations where the full simulation is less dense than the vanilla simulation by 2-$\sigma$ is highlighted by blue vertical bars. Bins below the noise threshold are not considered when measuring the differences between the two 1D profiles. The 1D profiles have been smoothed with a median box-car filter.}
        \label{fig:profiles}
        \end{figure*}

      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N = 10^5$). After binning the 1D profiles, we apply a boxcar smoothing technique. Each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs. This procedure reduces high-frequency noise and smooths the profiles, effectively removing the center of mass (see Fig.~\ref{fig:profiles}).

      With the smoothed 1D density profiles in hand, we treat the \textit{vanilla} simulations as the reference model and search for regions in the \textit{full} simulations that are significantly underdense compared to \textit{vanilla}, surpassing stochastic fluctuations. To accomplish this, we impose a signal-to-noise ratio threshold, $\mathcal{SNR}$. The signal is defined as the log of the counts per bin from the \textit{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution. We then compute a threshold for the number of counts in the \textit{vanilla} simulation, $N_v$, using the transcendental equation:
      \begin{equation}
          \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right) \sqrt{N_v}.
        \end{equation} \label{eq:density_threshold}
      By setting $\mathcal{SNR} = 5$, we solve for $N$ using \texttt{scipy.optimize.fsolve}, finding that $N$ must be greater than 7. After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \textit{vanilla} and full simulations:
      \begin{equation}
          \mathcal{R}_i = \log_{10}\left(\frac{N_{f,i}}{N_{v,i}}\right),
        \end{equation}
      where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the full simulation, and $N_{v,i}$ are the counts from the \textit{vanilla} simulation for each bin $i$. We then analyze the distribution of $\mathcal{R}_i$. If the differences between the density profiles are primarily due to stochastic processes of similar magnitude, this distribution should resemble a Gaussian as expected from the central limit theorem. Thus, we flag all regions where the density is underdense by more than two standard deviations, which should highlight regions whose underdensity is unlikely to be the result of the sum of stochastic processess but rather the passage of another globular cluster.

      This method, however, has its limitations, especially when detecting smaller gaps. As outlined by Erkal et al. (2015), a gap's size may not indicate a weak impact, but rather a more recent one. Additionally, since our streams have finite width, some gaps might be oblique to the stream axis. In such cases, marginalizing over $y'$ can dilute the gap's signal, making it harder to detect. This limitation is particularly evident in gaps caused by NGC2808, as discussed in the results. Therefore, this quantitative analysis serves as an \textit{aid} to visual inspection rather than a complete substitute for it.

@@

  We then transformed the stream to tail coordinates for the both the \textit{vanilla} and \textit{full} simulations in Fig.~\ref{fig:profiles}. As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations. Notice the \textit{Tiny Gap} labeled in the middle plot. This signal is not present in the 1D profile since it is too small in size and at an oblique angle, thus is lost when marginalizing. This process was followed for all 50 simulations to identify possible gaps. We find that small gaps are identifiable by eye rather well, where the 1D profiles aid in finding subtle yet extended underdensities. All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}. The profile method was useful in identifying the gaps in sampling 002, 006, 008, 029, 030.

>>>>

    \subsubsection*{Density Maps and Profiles}
      For each of the 50 simulations, we compare the final snapshots between the \texttt{vanilla} and \texttt{full} simulations using the tail coordinate system. We inspect these differences by generating 2D density maps of the streams. Additionally, we marginalize over the $y'$-coordinate to produce 1D density profiles along the $x'$-axis. These comparisons are shown in the results section (see Fig.~\ref{fig:profiles}).

      \begin{figure*}
        \centering
        \includegraphics[width=\linewidth, trim=20 0 15 0]{monte-carlo-009-pouliasis2017pii-GCNBody-2000-milisigma-5-noisefactor-20-boxcarindexlength-shifted-0.png}
        \caption{A comparison between the density maps and profiles of the \texttt{full} and \texttt{vanilla} simulations is presented. The locations where the full simulation is less dense than the vanilla simulation by 2-$\sigma$ is highlighted by blue vertical bars. Bins below the noise threshold are not considered when measuring the differences between the two 1D profiles. The 1D profiles have been smoothed with a median box-car filter.}
        \label{fig:profiles}
        \end{figure*}

      To construct the 1D density profiles, we first bin the data using the $\sqrt{N}$ rule, where $N$ is the number of data points ($N_p = 10^5$). After binning the 1D profiles, we apply a boxcar smoothing technique. Each central value is replaced by the median of the surrounding XX adjacent data points, corresponding to a smoothing length of approximately XX parsecs. This procedure reduces high-frequency noise and smooths the profiles. For instance, notice the absence of a high mass peak indicating for the center of mass in the bottom panel of Fig.~\ref{fig:profiles}.

      With the smoothed 1D density profiles in hand, we treat the \texttt{vanilla} simulations as the reference model and search for regions in the \texttt{full} simulations that are significantly underdense compared to \texttt{vanilla}, surpassing stochastic fluctuations. To accomplish this, we impose a signal-to-noise ratio threshold, $\mathcal{SNR}$. The signal is defined as the log of the counts per bin from the \texttt{vanilla} 1D density profile, and errors are propagated assuming a Poisson distribution. We then compute a threshold for the number of counts in the \texttt{vanilla} simulation, $N_v$, using the transcendental equation:
      \begin{equation}
          \mathcal{SNR} = \ln(10) \log_{10}\left(N_v\right) \sqrt{N_v}.
        \end{equation} \label{eq:density_threshold}
      By setting $\mathcal{SNR} = 5$, we solve for $N$ using \texttt{scipy.optimize.fsolve}, finding that $N$ must be greater than 7. After discarding insignificant bins (i.e., those with counts below the threshold), we compute the log ratio of the counts between the \texttt{vanilla} and full simulations:
      \begin{equation}
          \mathcal{R}_i = \log_{10}\left(\frac{N_{f,i}}{N_{v,i}}\right),
        \end{equation}
      where $\mathcal{R}_i$ is the log ratio, $N_{f,i}$ are the counts from the full simulation, and $N_{v,i}$ are the counts from the \texttt{vanilla} simulation for each bin $i$. We then analyze the distribution of $\mathcal{R}_i$. If the differences between the density profiles are primarily due to stochastic processes of similar magnitude, this distribution should resemble a Gaussian as expected from the central limit theorem. Thus, we flag all regions where the density is underdense by more than two standard deviations, which should highlight regions whose underdensity is unlikely to be the result of the sum of stochastic processess but rather the passage of another globular cluster.

      This method, however, has its limitations, especially when detecting smaller gaps. As outlined by Erkal et al. (2015), a gap's size may not indicate a weak impact, but rather a more recent one. Additionally, since our streams have finite width, some gaps are oblique with respect to the stream axis. In such cases, marginalizing over $y'$ erases the gap's signal, making it impossible to detect in a 1D profile. This limitation is particularly evident in gaps caused by NGC2808, as discussed in the results. Therefore, this quantitative analysis serves as an \textit{aid} to visual inspection rather than a complete substitute for it. This method helps particularly with large subtle gaps that the eye does not notice in the 2D maps.

@@

  We then transformed the stream to tail coordinates for the both the \texttt{vanilla} and \texttt{full} simulations in Fig.~\ref{fig:profiles}. As described in section~\ref{sec:gap_detection}, we locate regions on the stream where the full simulations are significantly less dense than the vanilla simulations. Notice the \textit{Tiny Gap} labeled in the middle plot. This signal is not present in the 1D profile since it is too small in size and at an oblique angle, thus is lost when marginalizing. This process was followed for all 50 simulations to identify possible gaps. We find that small gaps are identifiable by eye rather well, where the 1D profiles aid in finding subtle yet extended underdensities. All simulations and labeled gaps are presented in the Appendix~\ref{appendix:gallery_of_gaps}. The profile method was useful in identifying the gaps in sampling 002, 006, 008, 029, 030.
